---
title: "Salmon RMD"
author: "Iris Foxfoot"
date: "1/29/2021"
output: 
  html_document:
    theme: united
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(tsibble)
library(patchwork)
library(feasts)
```

## Willamette Falls Fish Passage Timeseries Summary {.tabset}

### Original Time Series

```{r}
#first read in the data
fish_passage <- read_csv(here("data", "willamette_fish_passage.csv")) %>% 
  clean_names()

#format dates and turn into tsibble  
fish_tsibble <- fish_passage %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  as_tsibble(key = NULL, index = date)

#select coho, jack coho, and steelhead
cjs_tsibble <- fish_tsibble %>% 
  select("coho", "jack_coho", "steelhead")

#Turn na to zero
cjs_tsibble[is.na(cjs_tsibble)] <- 0

#graph in ggplot, stored for late use in patchwork
coho_graph <- ggplot() + 
  geom_line(data = cjs_tsibble, 
            aes(x = date, y = coho),
            color = "deeppink3") +
  theme_minimal() +
  labs(x = "Year", y = "Coho")

steelhead_graph <- ggplot() +
  geom_line(data = cjs_tsibble, 
            aes(x = date, y = steelhead),
            color = "blue") +
  theme_minimal() +
  labs(x = "Year", y = "Steelhead")

jack_coho_graph <- ggplot() +
  geom_line(data = cjs_tsibble,
            aes(x = date, y = jack_coho),
            color = "darkgreen") +
  theme_minimal() +
  labs(x = "Year", y = "Jack Coho")

#make a patchwork graph and store it
fish_patchwork <- coho_graph/steelhead_graph/jack_coho_graph

#Call patchwork and add caption
fish_patchwork + plot_annotation(caption = "Time series of Salmon Species Passing Willamette Falls 2001-2010")
  
```

-All three species exhibit  clear seasonality 

-Steelhead are the most numerous fish of the three species to be observed

### Seasonplots
```{r}
#Using gg_season to create seasonplots and storing them for use in patchwork
coho_seasonplot <- gg_season(data = cjs_tsibble, y = coho, show.legend = FALSE) + #hiding all but one legend to reduce crowding
  theme_minimal() +
  labs(x = "Month",
       y = "Coho")


steelhead_seasonplot <- gg_season(data = cjs_tsibble, y = steelhead) +
  theme_minimal() +
  labs(x = "Month",
       y = "Steelhead")


jack_coho_seasonplot <- gg_season(data = cjs_tsibble, y = jack_coho, show.legend = FALSE) +
  theme_minimal() +
  labs(x = "Month",
       y = "Jack Coho")

#Patchwork season plots and storing
seasonplot_pw <- coho_seasonplot/steelhead_seasonplot/jack_coho_seasonplot 


#Calling out stored patchwork season plots
seasonplot_pw + plot_annotation(caption = "Season Plots of Salmon Species Passing Willamette Falls 2001-2010")

```

* The season for Coho seems to extend longer from 2008-2010

* Coho and Jack Coho have similar seasonality whereas the season for Steelheads corresponds with offseason for the Coho and Jack Coho

### Annual Counts by Species

(tab content here)
